Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers

### 2016-07-01 notes/paper/paper.tex
Updates to the paper.

The second module is the real-time activity classification on the unit.
This module reads in the reference(s) stored on the unit to be used as templates that the matched-filter will use for determining the individual\textquotesingle s activity.
Both of these modules use the same principles and hence the same algorithm kernels.
%
\section{The Algorithms}
%
\subsection{Activity Classification Method}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We use this method for identifying a user\textquotesingle s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter.

The matched filter is based on the cross correlation of the data $\textbf{y} = \{y\}_{k=0}^{N-1}$ with the reference signal $\textbf{x} = \{x\}_{k=0}^{M-1}$.
The cross correlation of $\textbf{x}$ with $\textbf{y}$ is defined as:
%
\begin{equation} \label{cross_correlation_eq}
(\textbf{x} \star \textbf{y}) = \left \{\sum_{i=0}^{M}x_{i} y_{i+k} \right \}_{k=0}^{N-M-1}
\end{equation}
%
which is simply a moving dot product.
We can then normalize \eqref{cross_correlation_eq} by the following:
%
\begin{equation} \label{norm_cross_correlation_eq}
\widehat{(\textbf{x} \star \textbf{y})}_k = \frac{(\textbf{x} \star \textbf{y})_k}{||\textbf{x}|| \ || \widetilde{\textbf{y}}_k || }, \quad k = 0,1,...,N-M-1
\end{equation}
%
where $|| \cdot ||$ is the usual $\ell_2$ norm and $\widetilde{\textbf{y}}_k = \{y_p\}_{p=k}^{M+k-1}$.
\eqref{norm_cross_correlation_eq} is now bounded by $[-1,1]$ and has any bias toward high energy removed.
We improve our computational efficiency by applying the convolution theorem to \eqref{cross_correlation_eq} and replacing it with:
%
\begin{equation} \label{conv_theorem}
(\textbf{x} \star \textbf{y}) = FFT^{-1}(FFT(\textbf{x}') FFT(\textbf{y})^*)
\end{equation}
%
where $\textbf{x}'$ is $\textbf{x}$ that has been zero padded to the length of $\textbf{y}$ with the added restriction that the length of $\textbf{x}$ is no more than the length of $\textbf{y}$ and $FFT$ is the Fast Fourier Transform algorithm.
We are most interested where our reference $\textbf{x}$ best matches our data $\textbf{y}$.
We use the resulting max correlation to define our metric for  our matched-filter output as:
%
\begin{equation} \label{matched_filter_eq}
MF_{\textbf{x}}(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x} \star \textbf{y})}_k \right \}
\end{equation}
%
If \eqref{matched_filter_eq} passes a set threshold, the corresponding signal is a match.
%
\subsection{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction with principal component analysis (PCA) \cite{bishop_2006}.
We assume that over a short period of time, a user\textquotesingle s limb movement is constrained mostly to a two-dimensional plane.
By identifying this plane of motion, we then project the acceleration data onto this plane and re-define our new axis accordingly.

If we let $\textbf{a} = \{a_i\}_{i=0}^{N}$ represent our demeaned acceleration data over a specified time window in three dimensions then the covariance matrix of our acceleration data is represented by the real-symmetric $3 \times 3$ matrix $Q = \textbf{a} \textbf{a}^T$.
The resulting unit length column eigenvectors of $Q$ are $V = (\textbf{v}^1,\textbf{v}^2,\textbf{v}^3)$ with their respective eigenvalues $\lambda^1 \geq \lambda^2 \geq \lambda^3$ of this matrix are orthonormal and point in the directions where the data varies most in descending order for each orthogonal direction.
% \ref{bishop_2006}
\begin{equation} \label{project_eq}
   \textbf{a}' = V^T \textbf{a}
\end{equation}
%
We now represent our acceleration data as $\textbf{a}'$ after projecting our acceleration data $\textbf{a}$ onto our rotated axis $V$ by \eqref{project_eq}. If the data is mostly dominate in the two primary dimensions, i.e. $\lambda^3 << \lambda^2$, then it is safe to assume that
%
\begin{equation} \label{magnitude_eq}
||(a^1,a^2,a^3)||^2 = ||(a'^1,a'^2,a'^3)||^2 \sim = (a'^1)^2 + (a'^2)^2
\end{equation}
$ \forall a \in \textbf{a}. $ \\
%
Equality comes from the orthonormality of $V$. The margin of error in the approximation is not important since we are only interested in processing in the two dimensions.
By now we have preserved as much information that is contained in our acceleration data as possible while also reducing our dimension from three to two.
By projecting our data onto the plane defined by the two most dominant eigenvectors, we effectively reduce our representation of our data to two dimensions.

To compute the matched-filter output for two dimensions, we generalize \eqref{matched_filter_eq} by augmenting our reference and data with both dimensions appropriately for normalization and adding together the cross correlations as:
%
\begin{equation} \label{cross_correlation_eq_2}
\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k = \frac{(\textbf{x}^1 \star \textbf{y}^1)_k + (\textbf{x}^2 \star \textbf{y}^2)_k}{||[ \textbf{x}^1, \textbf{x}^2 ]|| \ || [ \widetilde{\textbf{y}}_k^1, \widetilde{\textbf{y}}_k^2 ] || },
\end{equation}
%
for $ k = 0,1,...,N-M-1 $. \\
%
\begin{equation} \label{matched_filter_eq_2}
MF_{\textbf{x}}^2(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k \right \}
\end{equation}
%
This works because both dimensions are always in phase with each other for any given activity and therefore the match will happen in sync.
%
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of a signal.
The process consists of filtering to mitigate any possible aliasing distortion effects followed by a downsampling.
There is a trade off between performance and throughput when we decimate our signal, therefore, we use decimation solely for the purpose of reducing our processing cost.

For our application, we use a triangle filter prior to downsampling. This is achieved by convolving our data with the following weights:
%
\begin{equation} \label{triangle_filter_weights}
w_k^D =
\begin{cases}
  \frac{k+1}{D^2}, & \text{if }\ 0 \leq k < D \\
  \frac{2D - (k + 1)}{D^2}, & \text{if } D \leq k < 2D - 1 \\
  0, & \text{otherwise}
\end{cases}
\end{equation}
%
for a decimation value of $D$. Then the data is downsampled by $D$. This results in:
%
\begin{equation} \label{decimated_signal}
s_k^1 = \sum_{p=0}^{2N-1} w_{p+N(1-k)}^N s_{kN - (p+1)}^0 ?
\end{equation}
%
where $\{s_k^1\}_{k=0}^{M/D}$? is the decimated signal of $\{s_k^0\}_{k=0}^M$?.
\subsection{Training the Module}
We are able to identify an individual\textquotesingle s activity by matching the signal to a reference template.
For our application, we develop a reference signature for our matched filter directly from the person\textquotesingle s motion signature.
We do this by having the individual perform the specific activity over a specified amount of time and select the sub-interval of time that we find best represents the user's motion signature.
For a given activity, we attempt to find the representative periodic signal $s$ of our individual’s activity to be our template for our matched-filter.

If $\textbf{t} = \{t_k\}_{k=0}^{N_t}$ is our training set over $N_t$ data points, we define a collection of subsets $X$ of $\textbf{t}$:
%
\begin{equation} \label{X_subsets_of_training_eq}
X := \left \{ \{t_k\}_{k=p}^{p+M-1} | p=mI \right \}
\end{equation}
%
and $Y$ of $\textbf{t}$:
%
\begin{equation} \label{Y_subsets_of_training_eq}
Y := \left \{ \{t_k\}_{k=p}^{p+N-1} | p=nI \right \}
\end{equation}
%
where $M \leq N$, $I$ is a positive, fixed integer, and $(m,n) \in \mathbb{Z}^2$ where $(m,n)$ vary such that $0 \leq (m,n)I$ and $(m,n)I + (M,N) - 1 \leq N_t$. Then we can determine the representative signal $\textbf{s} \in X$ for our activity that satisfies the condition:
%
\begin{equation} \label{s_condition}
\sum_{y \in Y}MF^2_{\textbf{s}}(\textbf{y}) \geq \sum_{y \in Y}MF^2_{\textbf{x}}(\textbf{y}) \quad \forall \textbf{x} \in X
\end{equation}

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
\section{Results}
%
\subsection{Algorithm Performance}
%
The Kinetisense sensor was used for the algorithm performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit attached to the ankle of the user.
A separate instance of walking data has been used specifically for training the classifier.
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced at decimation 8. Figures (c) and (d) are the corresponding ROC curves produced from (a) and (b) respectively.}
  \label{fig:MF_performance}
\end{figure}
%
Figures \ref{fig:frac_walking_jogging} and \ref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in figures \ref{fig:roc_walking_jogging} and \ref{fig:roc_walking_bicycling} to illustrate the effectiveness of the matched-filter as a classifier.
Figures \ref{fig:frac_walking_bicycling} and \ref{fig:roc_walking_bicycling} show the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from bicycling.
%
%\FloatBarrier
%
\subsection{Throughput Performance}
%
To assess the real-time processing potential of the algorithms outlined in this paper, a throughput assessment was performed on the MSP432P401R microcontroller with varying parameters.
The Texas Instruments - MSP432P401R is a low power plus performance microcontroller that is based on the ARM 32-Bit Cortex-M4F processor.
Some of the notable features of the MSP432P410R is having a clock frequency of 48 MHz, 64 kB SRAM, 256 kB Flash Memory, and has an IEEE 754-compliant single-precision Floating-Point Unit (FPU) and Memory Protection Unit (MPU).
The single-precision processing capability on this unit provides us with a sufficiently large dynamic range necessary to process the algorithms that fixed-point processing cannot provide.
The 64 kB of SRAM provides sufficient space to process the results found in figure \ref{fig:timing} and
the 256 kB is more than enough flash memory to store the program and data references.
The clock frequency operates at 48 MHz which determines the execution speed.

@@

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds.}
   \label{fig:timing}
\end{figure}
%
Randomized data was used for the purpose of analyzing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows the execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process and operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even-powered factors, one can observe better throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even-valued time-lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than that the $50$Hz.
Figure \ref{fig:timing_refs} shows the execution time over a $4.0$ second time window at decimation $8$ while varying the sampling frequency and number of references. Because the pre-processing execution time does not change with the number of references, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\% when processing for a single reference.
%
\section{Conclusion}
We have demonstrated excellent activity classification capability of the matched-filter method using data obtained from a Kinetisense tri-axis accelerometer sensor.
We have shown that real-time processing can be achieved with timing results obtained from the MSP432 microcontroller.
Since FFT calculations dominate the processing time and work best with data that has radix 2 number of points, timing tests results demonstrate that radix 2 time-window lengths and frequency values result in faster processing times.
We increase our throughput with the use of data-compression techniques, namely decimation and dimensionality reduction, to reduce our computational requirements.
The references used in the matched-filter are obtained from the training module.
A training set is provided by the user through some specified activity and the best representative motion signature of the individual is extracted through the use of an instance-based learning algorithm.

>>>>

The second module is the real-time activity classification on the unit.
This module reads in the reference(s) stored on the unit to be used as templates that the matched-filter will use for determining the individual\textquotesingle s activity.
Both of these modules use the same principles and hence the same algorithm kernels.
%
\section{The Methodology}
%
\subsection{Activity Classification Method}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We use this method for identifying a user\textquotesingle s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter.

The matched filter is based on the cross correlation of the data $\textbf{y} = \{y\}_{k=0}^{N-1}$ with the reference signal $\textbf{x} = \{x\}_{k=0}^{M-1}$.
The cross correlation of $\textbf{x}$ with $\textbf{y}$ is defined as:
%
\begin{equation} \label{eq:cross_correlation}
(\textbf{x} \star \textbf{y}) = \left \{\sum_{i=0}^{M-1}x_{i} y_{i+k} \right \}_{k=0}^{N-M-1}
\end{equation}
%
which is simply a moving dot product.
We can then normalize \eqref{eq:cross_correlation} by the following:
%
\begin{equation} \label{eq:norm_cross_correlation}
\widehat{(\textbf{x} \star \textbf{y})}_k = \frac{(\textbf{x} \star \textbf{y})_k}{||\textbf{x}|| \ || \widetilde{\textbf{y}}_k || }, \quad k = 0,1,...,N-M-1
\end{equation}
%
where $|| \cdot ||$ is the usual $\ell_2$ norm and $\widetilde{\textbf{y}}_k = \{y_p\}_{p=k}^{M+k-1}$.
\eqref{eq:norm_cross_correlation} is now bounded by $[-1,1]$ and has any bias toward high energy removed.
We improve our computational efficiency by applying the convolution theorem to \eqref{eq:cross_correlation} and replacing it with:
%
\begin{equation} \label{eq:conv_theorem}
(\textbf{x} \star \textbf{y}) = FFT^{-1}(FFT(\textbf{x}') FFT(\textbf{y})^*)
\end{equation}
%
where $\textbf{x}'$ is $\textbf{x}$ that has been zero padded to the length of $\textbf{y}$ with the added restriction that the length of $\textbf{x}$ is no more than the length of $\textbf{y}$ and $FFT$ is the Fast Fourier Transform algorithm.
We are most interested where our reference $\textbf{x}$ best matches our data $\textbf{y}$.
We use the resulting max correlation to define our metric for  our matched-filter output as:
%
\begin{equation} \label{eq:matched_filter}
MF_{\textbf{x}}(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x} \star \textbf{y})}_k \right \}
\end{equation}
%
If \eqref{eq:matched_filter} passes a set threshold, the corresponding signal is a match.
%
\subsection{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction with principal component analysis (PCA) \cite{bishop_2006}.
We assume that over a short period of time, a user\textquotesingle s limb movement is constrained mostly to a two-dimensional plane.
By identifying this plane of motion, we then project the acceleration data onto this plane and re-define our new axis accordingly.

If we let $\textbf{a} = \{a_i\}_{i=0}^{N}$ represent our demeaned acceleration data over a specified time window in three dimensions then the covariance matrix of our acceleration data is represented by the real-symmetric $3 \times 3$ matrix $Q = \textbf{a} \textbf{a}^T$.
The resulting unit length column eigenvectors of $Q$ are $V = (\textbf{v}^1,\textbf{v}^2,\textbf{v}^3)$ with their respective eigenvalues $\lambda^1 \geq \lambda^2 \geq \lambda^3$ of this matrix are orthonormal and point in the directions where the data varies most in descending order for each orthogonal direction.
% \ref{bishop_2006}
\begin{equation} \label{eq:project}
   \textbf{a}' = V^T \textbf{a}
\end{equation}
%
We now represent our acceleration data as $\textbf{a}'$ after projecting our acceleration data $\textbf{a}$ onto our rotated axis $V$ by \eqref{eq:project}. If the data is mostly dominate in the two primary dimensions, i.e. $\lambda^3 << \lambda^2$, then it is safe to assume that
%
\begin{equation} \label{eq:magnitude}
||(a^1,a^2,a^3)||^2 = ||(a'^1,a'^2,a'^3)||^2 \sim = (a'^1)^2 + (a'^2)^2
\end{equation}
$ \forall a \in \textbf{a}. $ \\
%
Equality comes from the orthonormality of $V$. The margin of error in the approximation is not important since we are only interested in processing in the two dimensions.
By now we have preserved as much information that is contained in our acceleration data as possible while also reducing our dimension from three to two.
By projecting our data onto the plane defined by the two most dominant eigenvectors, we effectively reduce our representation of our data to two dimensions.

To compute the matched-filter output for two dimensions, we generalize \eqref{eq:matched_filter} by augmenting our reference and data with both dimensions appropriately for normalization and adding together the cross correlations as:
%
\begin{equation} \label{eq:cross_correlation_2}
\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k = \frac{(\textbf{x}^1 \star \textbf{y}^1)_k + (\textbf{x}^2 \star \textbf{y}^2)_k}{||[ \textbf{x}^1, \textbf{x}^2 ]|| \ || [ \widetilde{\textbf{y}}_k^1, \widetilde{\textbf{y}}_k^2 ] || },
\end{equation}
%
for $ k = 0,1,...,N-M-1 $. \\
%
\begin{equation} \label{eq:matched_filter_2}
MF_{\textbf{x}}^2(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k \right \}
\end{equation}
%
This works because both dimensions are always in phase with each other for any given activity and therefore the match will happen in sync.
%
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of a signal.
The process consists of filtering to mitigate any possible aliasing distortion effects followed by a downsampling.
There is a trade off between performance and throughput when we decimate our signal, therefore, we use decimation solely for the purpose of reducing our processing cost.

For our application, we use a triangle filter prior to downsampling. This is achieved by applying the following weights:
%
\begin{equation} \label{eq:triangle_filter_weights}
w_k^D =
\begin{cases}
  \frac{D+k}{D^2}, & \text{if }\ -D < k \leq 0 \\
  \frac{D-k}{D^2}, & \text{if } D < k < D \\
  0, & \text{otherwise}
\end{cases}
\end{equation}
%
and convolve it with our data to get
%
\begin{equation} \label{eq:convolve}
s'_k = \sum_{i=0}^{N} s_i w_{k-i}
\end{equation}
%
followed by a downsample
%
\begin{equation} \label{eq:downsample}
s''_k = s'_{(k+1)D-1}, \quad k = 0, 1, 2, ..., M/D.
\end{equation}
%
One can reduce the processing cost by combining \eqref{eq:convolve} and \eqref{eq:downsample} together in such a way that we only process the points that we keep.
%
\subsection{Training the Module}
We are able to identify an individual\textquotesingle s activity by matching the signal to a reference template.
For our application, we develop a reference signature for our matched filter directly from the person\textquotesingle s motion signature.
We do this by having the individual perform the specific activity over a specified amount of time and select the sub-interval of time that we find best represents the user's motion signature.
For a given activity, we attempt to find the representative periodic signal $s$ of our individual’s activity to be our template for our matched-filter.

If $\textbf{t} = \{t_k\}_{k=0}^{N_t}$ is our training set over $N_t$ data points, we define a collection of subsets $X$ of $\textbf{t}$:
%
\begin{equation} \label{eq:X_subsets_of_training_eq}
X := \left \{ \{t_k\}_{k=p}^{p+M-1} | p=mI \right \}
\end{equation}
%
and $Y$ of $\textbf{t}$:
%
\begin{equation} \label{eq:Y_subsets_of_training_eq}
Y := \left \{ \{t_k\}_{k=p}^{p+N-1} | p=nI \right \}
\end{equation}
%
where $M \leq N$, $I$ is a positive, fixed integer, and $(m,n) \in \mathbb{Z}^2$ where $(m,n)$ vary such that $0 \leq (m,n)I$ and $(m,n)I + (M,N) - 1 \leq N_t$. Then we can determine the representative signal $\textbf{s} \in X$ for our activity that satisfies the condition:
%
\begin{equation} \label{eq:s_condition}
\sum_{y \in Y}MF^2_{\textbf{s}}(\textbf{y}) \geq \sum_{y \in Y}MF^2_{\textbf{x}}(\textbf{y}) \quad \forall \textbf{x} \in X
\end{equation}

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
\section{Performance and Throughput}
%
\subsection{Algorithm Performance}
%
The Kinetisense sensor was used for the algorithm performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit attached to the ankle of the user.
A separate instance of walking data has been used specifically for training the classifier.
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced at decimation 8. Figures (c) and (d) are the corresponding ROC curves produced from (a) and (b) respectively.}
  \label{fig:MF_performance}
\end{figure}
%
Figures \ref{fig:frac_walking_jogging} and \ref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in figures \ref{fig:roc_walking_jogging} and \ref{fig:roc_walking_bicycling} to illustrate the effectiveness of the matched-filter as a classifier.
Figures \ref{fig:frac_walking_bicycling} and \ref{fig:roc_walking_bicycling} show the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from bicycling.
%
%\FloatBarrier
%
\subsection{Throughput Performance}
%
To assess the real-time processing potential of the algorithms outlined in this paper, a throughput assessment was performed on the MSP432P401R microcontroller with varying parameters.
The Texas Instruments - MSP432P401R is a low power plus performance microcontroller that is based on the ARM 32-Bit Cortex-M4F processor.
Some of the notable features of the MSP432P410R is having a clock frequency of 48 MHz, 64 kB SRAM, 256 kB Flash Memory, and has an IEEE 754-compliant single-precision Floating-Point Unit (FPU) and Memory Protection Unit (MPU).
The single-precision processing capability on this unit provides us with dynamic range necessary to process the algorithms that fixed-point processing cannot provide.
The 64 kB of SRAM provides sufficient space to process the results found in figure \ref{fig:timing} and
the 256 kB is more than enough flash memory to store the program and data references.
The clock frequency operates at 48 MHz which determines the execution speed.

@@

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds with timings measured at sampling frequencies of $16, 32, 62, 64, 75, 100,$ and $128$ Hz.}
   \label{fig:timing}
\end{figure}
%
Randomized data was used for the purpose of analyzing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows the execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process and operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even-powered factors, one expects best throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even-valued, non-radix 2, time-lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than that the $50$Hz.
Figure \ref{fig:timing_refs} shows the execution time over a $4.0$ second time window at decimation $8$ while varying the sampling frequency and number of references. Because the pre-processing execution time does not change with the number of references, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\% when processing for a single reference.
%
\section{Conclusion}
We have demonstrated excellent activity classification capability of the matched-filter method using data obtained from a Kinetisense tri-axis accelerometer sensor.
We have shown that real-time processing can be achieved with timing results obtained from the MSP432 microcontroller.
Timing results have demonstrated that radix 2 time-window lengths and frequency values result in faster processing times as expected given that the FFT dominates the procesing time.
We increase our throughput with the use of data-compression techniques, namely decimation and dimensionality reduction, to reduce our computational requirements.
The references used in the matched-filter are obtained from the training module.
A training set is provided by the user through some specified activity and the best representative motion signature of the individual is extracted through the use of an instance-based learning algorithm.
